Reproducible Research Peer Assessment 1
=========================================

##1. Loading and preprocessing the data

The data for the assessment is loaded into R using the read.csv() function
```{r echo=FALSE}
library(lubridate)
library(ggplot2)
library(lattice)
setwd("C:/Users//Shyam/Videos/Coursera - Data Science Specialization/Reproducable Research/Peer Assessment 1")
```
```{r}
data <- read.csv("activity.csv",header=TRUE)
str(data)
summary(data)
```


##2. Mean total number of steps taken per day

* Calculate Total Steps excluding the NA Values in the data set 
```{r}
total_steps_per_day <- aggregate(data$steps,by = list(data$date),sum,na.rm=TRUE)
names(total_steps_per_day) = c("Date","Total")
head(total_steps_per_day)
```

* Display the histogram and compute Mean and Median 
```{r}
hist(total_steps_per_day$Total, col= "blue",xlab = "Total No Of Steps",main = "Histogram of Total No of Steps")
#Calculate Mean and Median
mean <- mean(total_steps_per_day$Total)
median <- median(total_steps_per_day$Total)
```

Mean :  `r round(mean,2)` 

Median :  `r median` 


##3. Average daily activity pattern

* Compute Average no of steps per interval
```{r}
average<- aggregate(data$steps, by = list(data$interval),mean,na.rm=TRUE)
names(average) = c("Intervals","AverageSteps")
head(average)
```

* Time Series Plot of 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
```{r}
plot(x = average$Intervals, y=average$AverageSteps, type="l",col="red",ylab="Average No of Steps",
     xlab="Intervals", main="Time Series Plot")
maximum<-average[which.max(average$AverageSteps),1]
```

5 Minute interval that has maximum number of steps : `r maximum`


##4. Imputing missing values

* Get the total no of NA Values
```{r}
count_of_na <- sum(is.na(data$steps))
na_positions <- which(is.na(data$steps))
```
The no of NAs : `r count_of_na`  

* Strategy for filling the NA Values. Here mean no of steps is used.
```{r}
vector_of_means <- rep(mean(data$steps, na.rm=TRUE), length(na_positions))
data[na_positions, "steps"] <- vector_of_means
head(data)
```

* Computing total_steps with the NAs populated.
```{r}
total_steps <- aggregate(data$steps,by = list(data$date),sum,na.rm=TRUE)
names(total_steps) = c("Date","Total")
head(total_steps)
hist(total_steps$Total, col= "blue",xlab = "Total No Of Steps",main = "Histogram of Total No of Steps")
```

* Computing Mean and Median
```{r}
new_mean <- mean(total_steps$Total)
new_median <- median(total_steps$Total)
```

Mean : `r round(new_mean,3)`

Median : `r round(new_median,3)`


##5. Differences in activity patterns between weekdays and weekends

* Getting Days of Week and populating the appropriate value
```{r}
data$day_of_week<- ifelse(weekdays(as.Date(data$date)) %in% c("Saturday", "Sunday"),'weekend','weekday')                                                
```

* Computing average number of steps taken, averaged across all weekday days or weekend days
```{r}
average<- aggregate(data$steps, by=list(data$interval,data$day_of_week),mean)
names(average)=c("Interval","Day","Mean")
xyplot(average$Mean~average$Interval|average$Day,type='l',layout=c(1,2),
       xlab="Intervals",ylab="No of Steps Taken",
       main ="Difference in Activity patterns between Weekdays & Weekends")
```
